---
title: "Homework - week 1"
author: "Mateus Silvestrin"
format: pdf
---

```{r}
#| message: false
library(dplyr)
library(tibble)
library(ggplot2)

theme_set(theme_bw(14))
```


### 1. Suppose the globe tossing data (Lecture 2, Chapter 2) had turned out to be 3 water and 11 land. Construct the posterior distribution.

To do this, we only need to change the call to `dbinom`. In the original example we had 6 water draws, so that went into the `x` argument of the function. Also, the original example had a total of nine draws, and that went into the `size` argument. We change that value accordingly.
```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prob_p <- rep( 1 , 1000 )
prob_data <- dbinom( 3 , size=14 , prob=p_grid ) # changes are here
posterior <- prob_data * prob_p
posterior <- posterior / sum(posterior)
```


```{r}
tibble("Probability of water"  = p_grid,
           "Posterior probability" = posterior) |>
  ggplot(aes(`Probability of water`,`Posterior probability`))+
  geom_line(colour = "lightskyblue",
            linewidth = 2)+
  labs(title = "Posterior distribution for the globe model", 
       caption = "Observed values = 3W, 11L.")
```


### 2. Using the posterior distribution from 1, compute the posterior predictive distribution for the next 5 tosses of the same globe. I recommend you use the sampling method.

First, sample from the posterior distribution:

```{r}
#| results: hold
samples <- sample(p_grid, prob=posterior, size=1e4, replace=TRUE)
"Summary of samples:"
summary(samples)
```

Simulate sampling from the samples of the posterior distribution:
```{r}
#| results: hold
w <- rbinom( 1e4, size=5, prob=samples )
"Summary of samples:"
table(w)
```


Visualisation:
```{r}
tibble(w = w) |>
  group_by(w) |>
  count() |>
  ungroup() |>
  mutate(p = n/1e4) |>
  ggplot(aes(w,p))+
  geom_col(fill = "lightskyblue", 
           colour = "gray20",
           width = .3)+
  scale_y_continuous(expand = expansion(c(0,.15)))+
  labs(title = "Predictive posterior distribution",
       subtitle = "(Five tosses)",
       x = "Number of observed water")
```

