---
title: "Homework - week 3"
author: "Mateus Silvestrin"
format: 
  pdf:
    documentclass: scrreprt
    mainfont: "FreeSerif"
    code-block-bg: true
---

```{r}
#| message: false
library(data.table)
library(ggplot2)
library(dagitty)
library(rethinking)

data("foxes")
setDT(foxes)
theme_set(theme_bw(14))
```

```{r}
#| warning: false
foxes |> 
  melt("group") |> 
  ggplot(aes(value))+
  geom_histogram()+
  facet_grid(variable~.)
```

Standardizing variables.
```{r}
foxes_std <- foxes
foxes_std[, 2:5 := lapply(.SD, standardize), .SDcols= -"group"]
```


## 1. Total causal influence of A on F.
```{r}
sim_AF <- 
  function(A, bA){
    N <- length(A)
    Fd <- rnorm(N, bA*A, .5)
    data.table(A,Fd)
  }

sim_AF(A = runif(1e2,-2,2), bA = .5) |> 
  ggplot(aes(A, Fd))+
  geom_point()
```

```{r}
m_AF <- 
  alist(
    Fd <- dnorm(mu, sigma),
    mu <- bA*A,
    # priors
    bA ~ dunif(0,2),
    sigma ~ dunif(0,2)
)

fit_AF <- 
  quap(m_AF,data = list(A = foxes_std$area,
                        Fd = foxes_std$avgfood))

precis(fit_AF)
```

```{r}
post_AF <- extract.samples(fit_AF)

post_AF |> 
  ggplot(aes(bA))+
  geom_density(fill = "darkslategray2", alpha = .5)+
  scale_y_continuous(breaks = seq(0,10,2),limits = c(0,10),
                     expand = expansion(c(.01,.05)))+
  labs(x = "Effect of Area on Food (standardized)",
       subtitle = "Posterior distribution")
```

Increasing Area leads to more Food.

## 2. Total causal effect of Food on Weight

Intervention to estimate total causal effect of F on W. It implies a DAG where area does not exist. Also, since F -> G -> W is a pipe (or moderation), to calculate the total effect of F we should not include G in the model.

```{r}
m_FW <- 
  alist(
    # F -> W <- G
    W <- dnorm(mu, sigma),
    mu <- bFd*Fd ,
    # priors
    bFd ~ dunif(-1,4),
    sigma ~ dunif(0,2)
)

fit_FW <- 
  quap(m_FW,data = list(W = foxes_std$weight,
                        Fd = foxes_std$avgfood))
```

```{r}
post_FW <- extract.samples(fit_FW)

post_FW |> 
  ggplot(aes(bFd))+
  geom_density(fill = "darkslategray2", alpha = .5)+
  scale_y_continuous(breaks = seq(0,4,1),limits = c(0,4.5),
                     expand = expansion(c(.01,.05)))+
  labs(x = "Total effect of Food on Weight (standardized)",
       subtitle = "Posterior distribution")
```

### Now, the intervention.

```{r}
Fd_seq <- data.table(Fd = seq(-2,2,.25))
sim_FW <- sim(fit_FW, data = Fd_seq, vars = c("W"))
```

```{r}
sim_FW <- setnames(as.data.table(sim_FW), paste0("V",1:17), as.character(Fd_seq$Fd))
#lapply(as.data.table) |> 
#rbindlist(idcol = "counterfactual") |> 

sim_FW[, counterfactual := "W"]
sim_FW <-  
  sim_FW |> 
  melt("counterfactual", variable.name = "Food", variable.factor = F)
sim_FW[,Food := as.numeric(Food)]

sim_FW[,
        .(M = mean(value),
          ymin = PI(value)[1],
          ymax = PI(value)[2]),
        by = .(counterfactual, Food)]|> 
  ggplot(aes(Food, ymin = ymin, ymax = ymax))+
  geom_ribbon(alpha = .4)+
  geom_line(aes(y = M))+
  labs(y = "W")
  #+
  #facet_wrap(counterfactual~.)
```

```{r}
sim_FW_wide <- 
  dcast(cbind(sim_FW[,-"Food"],sim_FW[,.(n = 1:.N),by = Food]),
        counterfactual + n ~ Food)
sim_FW_wide[, .(effect = `1`-`0`)] |> 
  ggplot(aes(effect)) +
  geom_density(fill = "firebrick", alpha = .5)+
 scale_y_continuous(breaks = seq(0,.3,.1),limits = c(0,.3),
                     expand = expansion(c(.01,.05)))+
  labs(x = "Total causal effect of 1sd Food increase in Weight",
       subtitle = "Contrast")

sim_FW_wide[, .(effect = `-1`-`0`)] |> 
  ggplot(aes(effect)) +
  geom_density(fill = "firebrick", alpha = .5)+
  scale_y_continuous(breaks = seq(0,.3,.1),limits = c(0,.3),
                     expand = expansion(c(.01,.05)))+
  labs(x = "Total causal effect of 1sd Food decrease in Weight",
       subtitle = "Contrast")
```


## 3. Direct effect of adding Food on Weight
```{r}
m_FGW <- 
  alist(
    # F -> W <- G
    W <- dnorm(mu, sigma),
    mu <- bFd*Fd + bG*G,
    # priors
    bFd ~ dunif(-1,4),
    bG ~ dunif(-4,0), # the bigger the group, lesser the weight
    sigma ~ dunif(0,2)
)

fit_FGW <- 
  quap(m_FGW,data = list(W = foxes_std$weight,
                        Fd = foxes_std$avgfood,
                        G = foxes_std$groupsize))
```

```{r}
post_FGW <- extract.samples(fit_FGW)

post_FGW |> 
  ggplot(aes(bFd))+
  geom_density(fill = "darkslategray2", alpha = .5)+
  scale_y_continuous(#breaks = seq(0,4,1), limits = c(0,4.5),
                     expand = expansion(c(.01,.05)))+
  labs(x = "Direct effect of Food on Weight (standardized)",
       subtitle = "Posterior distribution")
```

```{r}
sim_FGW <- sim(fit_FGW,
               data = list(Fd = c(0,1)),
               vars = c("W","G"))

data.table(contrast_1sd = sim_FGW$W[,2] - sim_FGW$W[,1]) |> 
  ggplot(aes(contrast_1sd)) +
  geom_density(fill = "firebrick", alpha = .5)+
 scale_y_continuous(#breaks = seq(0,.3,.1),limits = c(0,.32),
                     expand = expansion(c(.01,.05)))+
  labs(x = "Direct causal effect of 1sd Food increase in Weight",
       subtitle = "Contrast")
```

```{r}
sim_FGW2 <- sim(fit_FGW,
               data = list(G = c(0,1)),
               vars = c("W","Fd"))

data.table(contrast_1sd = sim_FGW2$W[,2] - sim_FGW2$W[,1]) |> 
  ggplot(aes(contrast_1sd)) +
  geom_density(fill = "firebrick", alpha = .5)+
 scale_y_continuous(#breaks = seq(0,.3,.1),limits = c(0,.35),
                     expand = expansion(c(.01,.05)))+
  labs(x = "Direct causal effect of 1sd Group Size increase on Weight",
       subtitle = "Contrast")
```

Strikingly, more food does not seems to lead to higher weight in these foxes. However, more food leads to higher group sizes, which causes the average weight to go down. 